cardinalR: Generating interesting high-dimensional data structures

### 2025-05-07 paper-cardinalR.Rmd
Removed the subsection of application
Try to fix the issues of compiling the Latex code

where:
$\epsilon \sim \mathcal{U}(0, 0.1)$ adds local jitter
$s\_i \in \[0.5, 2]$ is randomly sampled for each branch.

@@

Branches $3$ to $k$** are added iteratively. Each new branch begins at a randomly chosen point within a restricted horizontal band: $X\_1 \in \[-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X\_1\$:$X\_1 \in \[x\_{\text{start}}, x\_{\text{start}} + 1]$, and has the structure:

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let $X\_1 \sim \mathcal{U}(\text{range}\[1], \text{range}\[2])$ be the independent variable. A raw polynomial basis of degree 2 is used to form $X\_2 = X\_1 - X\_1^2 + \varepsilon\_2$, where $\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.

@@

The shape is generated using a third-degree raw polynomial basis expansion. Let $X\_1 \sim \mathcal{U}(\text{range}\[1], \text{range}\[2])$ be the base input. A cubic transformation with vertical jitter is used to define $X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.

@@

Each point is placed within a regular hexagon in the plane $(x\_1, x\_2)$, using a randomly chosen hexagon sector angle $\theta \in \left{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3} \right}$ and a uniformly random radial scaling factor:

@@

Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T\_0 \in \[0, 1]^p$ drawn from a uniform distribution:

@@

$$
X_j = s_j \cdot \theta + \varepsilon_j$,\text{ where }$\varepsilon_j \sim \mathcal{N}(0, 0.5^2)$,\text{ and }$s_j$\text{ is a random scale}.
$$

@@

$$
X_j \sim \mathcal{N}(0, 0.05^2)$,\text{ for }$j = 5, \ldots, p.
$$

@@

## Assessing the performance of dimension reduction on different geometric structures in high-dimensions

>>>>

where:
$\epsilon \sim \mathcal{U}(0, 0.1)$ adds local jitter,

$s\_i \in [0.5, 2]$ is randomly sampled for each branch.

@@

Branches $3$ to $k$** are added iteratively. Each new branch begins at a randomly chosen point within a restricted horizontal band: $X\_1 \in [-0.15, 0.15]$ (to ensure connectivity with earlier branches), spans a unit-length interval on $X\_1$:$X\_1 \in [x\_{\text{start}}, x\_{\text{start}} + 1]$, and has the structure:

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let $X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the independent variable. A raw polynomial basis of degree 2 is used to form $X\_2 = X\_1 - X\_1^2 + \varepsilon\_2$, where $\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.

@@

The shape is generated using a third-degree raw polynomial basis expansion. Let $X\_1 \sim \mathcal{U}(\text{range}[1], \text{range}[2])$ be the base input. A cubic transformation with vertical jitter is used to define $X\_2 = X\_1 + X\_1^2 - X\_1^3 + \varepsilon\_2\$, where \$\varepsilon\_2 \sim \mathcal{U}(0, 0.5)$.

@@

Each point is placed within a regular hexagon in the plane $(x\_1, x\_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor:

@@

Let $x\_1, x\_2, \dots, x\_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T\_0 \in [0, 1]^p$ drawn from a uniform distribution:

@@

$$
X_j = s_j \cdot \theta + \varepsilon_j,\text{ where }\varepsilon_j \sim \mathcal{N}(0, 0.5^2),\text{ and }s_j\text{ is a random scale}.
$$

@@

$$
X_j \sim \mathcal{N}(0, 0.05^2),\text{ for }j = 5, \ldots, p.
$$

@@

<!-- ## Assessing the performance of dimension reduction on different geometric structures in high-dimensions -->
